Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-04-14 paper/arm.tex
intro

\section{Efficient Implementation on ARMv8}
\label{sec:arm}

We discuss \armeight{} using \emph{external global completion} (\EGC{})
\cite{alglave-git-alternate} %,armed-cats},
\cite[\textsection B2.3.6]{arm-reference-manual}
which is very close to our model.

\subsection{Downgraded Reads (\xDGR)}
\label{sec:downgrade}

\begin{example}
  \label{ex:dgr1}
  The following example is from \citeauthor{anton} \cite{anton}.  It is
  allowed by \armeight{}, but disallowed by \refdef{def:pomsets-ra}.  The
  coherence order between the writes can be witnessed by a separate thread,
  which we have elided.
  \begin{gather*}
    \PW{x}{2}\SEMI 
    \PR[\mRA]{x}{r}\SEMI
    \PW{y}{1} \PAR
    \PW{y}{2}\SEMI
    \PW[\mRA]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DW{x}{2}}{}
        \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
        \event{c}{\DW{y}{1}}{right=of b}
        \event{d}{\DW{y}{2}}{right=2.5em of c}
        \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
        \rf{a}{b}
        % \sync{b}{c}
        \wk{c}{d}
        \sync{d}{e}
        \wk[out=-165,in=-15]{e}{a}
      \end{tikzinline}}  
  \end{gather*}

  More generally, this can be understood as a compiler optimization that
  downgrades a read from $\mRA$ to $\mRLX$ when it can be fulfilled by a
  relaxed write in the same thread.
\end{example}

To model such \emph{downgraded reads}, we use the uninterpreted symbols
$\Dx{\aLoc}$.

\begin{definition}
  \label{def:DS}
  % Let $[\FALSE/\D]$ be the substitution that performs $[\FALSE/\Dx{\aLoc}]$
  % for every $\aLoc$.
  Let $[\aForm/\Dx{*}]$ substitute $\aForm$ for every $\Dx{\bLoc}$.\\
  Let formula $\DL{\aLoc}{\amode}$ and substitution $\DS{\aLoc}{\amode}$ be defined:
  \begin{align*}
    \DL{\aLoc}{\mRLX}&=\TRUE
    &\DS{\aLoc}{\mRLX}&=[\TRUE/\Dx{\aLoc}] 
    \\
    \DL{\aLoc}{\mRA}&=\Dx{\aLoc}%\land\D{}
    &\DS{\aLoc}{\mRA}&=[\FALSE/\D]
    \\
    \DL{\aLoc}{\mSC}&=\Dx{\aLoc}%\land\D{}
    &\DS{\aLoc}{\mSC}&=[\FALSE/\D]
  \end{align*}
\end{definition}
\begin{definition}[\xCO/\xRASC/\xDGR]
  \label{def:pomsets-down}
  Update \refdef{def:pomsets-ra} to: % (\ref{L4} unchanged):
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies
    \begin{math}
      \bForm
      \DS{\aLoc}{\amode}
      [(\Qw{\aLoc}\land\aExp{=}\aVal)/\Qw{\aLoc}],
    \end{math}
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      \bForm
      \DS{\aLoc}{\amode}
      [\FALSE/\QS{\aLoc}{\amode}],
    \end{math}
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies
    \begin{math}
      \aVal{=}\aReg\limplies\bForm,
    \end{math}
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      \DL{\aLoc}{\amode}\land \bForm[\FALSE/\QL{\aLoc}{\amode}].
    \end{math}
  \end{enumerate}
\end{definition}
Load actions that require downgrading introduce $\Dx{\aLoc}$.
Relaxed stores on $\aLoc$ substitute true for $\Dx{\aLoc}$, whereas
synchronizing stores substitute false for $\Dx{\aLoc}$.

\begin{example}
  Revisiting \refex{ex:dgr1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of a}
          % \xform{x1i}{\lnot\Qw{x}\land\bForm[\TRUE/\Dx{x}]}{below=of x1d}
          \xo{a}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR[\mRA]{x}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\DR[\mRA]{x}{2}}{}
          % \xform{x1d}{2{=}r\limplies\bForm}{below=of b}
          \xform{x1i}{\Dx{x} \land\bForm[\FALSE/\Qr{x}]}{below=of b} %x1d}
          % \xo{b}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{\DW{y}{1}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Associating right:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{right=.7em of a}
          % \xform{x1i}{\lnot\Qw{x}\land\bForm[\TRUE/\Dx{x}]}{below=of x1d}
          \xos{a}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR[\mRA]{x}{r}
      \SEMI
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\DR[\mRA]{x}{2}}{}
          \event{c}{\Dx{x}\mid\DW{y}{1}}{right=.7em of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing, we have, as desired:
  \begin{align*}
    \begin{gathered}[t]
      \PW{x}{2}
      \SEMI
      \PR[\mRA]{x}{r}
      \SEMI
      \PW{y}{1}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{a}{\DW{x}{2}}{}
          \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
          \event{c}{\DW{y}{1}}{right=of b}
          \rf{a}{b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
\end{example}

% The following execution is allowed both by \armeight{} and \tso{}, but
% disallowed by \refdef{def:pomsets-ra}.
% \begin{gather*}
%   \PW{x}{2}\SEMI 
%   \PR[\mRA]{x}{r}\SEMI
%   \PR{y}{s} \PAR
%   \PW{y}{2}\SEMI
%   \PW[\mRA]{x}{1}
%   \\
%   \hbox{\begin{tikzinline}[node distance=1.5em]
%     \event{a}{\DW{x}{2}}{}
%     \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
%     \event{c}{\DR{y}{0}}{right=of b}
%     \event{d}{\DW{y}{2}}{right=2.5em of c}
%     \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
%     \rf{a}{b}
%     \sync{b}{c}
%     \wk{c}{d}
%     \sync{d}{e}
%     \wk[out=-165,in=-15]{e}{a}
%   \end{tikzinline}}
% \end{gather*}

\begin{example}
  One might worry that our model is too permissive for $\mSC$ access, but
  \armeight{} itself allows some very counterintuitive results for $\mSC$
  access.  In the following execution we elide the initializing write
  $\DWP{y}{0}$.
  \begin{gather*}
    \IF{\PR{x}{}}\THEN \PW{x}{2} \FI\SEMI
    \PR[\mSC]{x}{r}\SEMI
    \PR[\mSC]{y}{s} \PAR
    \PW[\mSC]{y}{2}\SEMI
    \PW[\mSC]{x}{1}
    \\
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{a}{\DR{x}{1}}{}
        \event{b}{\DW{x}{2}}{right=of a}
        \scevent{c}{\DR[\mSC]{x}{2}}{right=of b}
        \scevent{d}{\DR[\mSC]{y}{0}}{right=of c}
        \scevent{e}{\DW[\mSC]{y}{2}}{right=2.5em of d}
        \scevent{f}{\DW[\mSC]{x}{1}}{right=of e}
        \po{a}{b}
        \rf{b}{c}
        % \sync{c}{d}
        \sync{e}{f}
        \wk{d}{e}
        \rf[out=-165,in=-15]{f}{a}
      \end{tikzinline}}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP[\mSC]{x}{2}\xsync\DRP{y}{0}$, because $\DRP[\mSC]{x}{2}$ is fulfilled
  by a relaxed write in the same thread.
\end{example}

\subsection{Removing Read-Read dependencies (\xRRD)}
\label{sec:rrd}

\begin{example}
  \label{ex:rrd1}
  The following execution is allowed by \armeight{}, but disallowed by
  \refdef{def:pomsets-ra}.  
  \begin{gather*}
    \begin{gathered}
      x\GETS0\SEMI x\GETS 1\SEMI y^\mRA\GETS1
      \PAR
      r\GETS y\SEMI \IF{r}\THEN s\GETS x\FI
      \\[-.4ex]
      \hbox{\begin{tikzinline}[node distance=1.5em]
          \event{wx0}{\DW{x}{0}}{}
          \event{wx1}{\DW{x}{1}}{right=of wx0}
          \raevent{wy1}{\DW[\mRA]{y}{1}}{right=of wx1}
          \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
          \event{rx0}{\DR{x}{0}}{right=of ry1}
          \sync{wx1}{wy1}
          \po{ry1}{rx0}
          \rf{wy1}{ry1}
          \wk[in=-15,out=-165]{rx0}{wx1}
          \wk{wx0}{wx1}
        \end{tikzinline}}
    \end{gathered}
  \end{gather*}
  Under \EGC{}, this is explained by dropping the order
  $\DRP{y}{1}\xpo\DRP{x}{0}$, because \armeight{} does not include control
  dependencies between reads in the locally-ordered-before relation.
  \begin{gather*}
    \hbox{\begin{tikzinline}[node distance=1.5em]
        \event{wx0}{\DW{x}{0}}{}
        \event{wx1}{\DW{x}{1}}{right=of wx0}
        \raevent{wy1}{\DWRel{y}{1}}{right=of wx1}
        \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
        \event{rx0}{\DR{x}{0}}{right=of ry1}
        \sync{wx1}{wy1}
        % \po{ry1}{rx0}
        \rf{wy1}{ry1}
        \wk[in=-15,out=-165]{rx0}{wx1}
        \wk{wx0}{wx1}
      \end{tikzinline}}  
  \end{gather*}
\end{example}

Since we do not distinguish control dependencies from other dependencies, we
are forced to drop all dependencies between reads.  In order to do so, we use
the uninterpreted symbol $\RW$.

\begin{definition}[\xRRD]
  \label{def:pomsets-rr}
  Update \refdef{def:pomsets-trans} to:
  % \begin{enumerate}
  % \item[\ref{T3})]
  %   $\labelingForm(\aEv)$ implies
  %   $\labelingForm_1(\aEv)[\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
  %   \\
  %   $\labelingForm(\aEv)$ implies
  %   $\labelingForm_1(\aEv)[\FALSE/\RW]$ otherwise.
  % \end{enumerate}
  \begin{enumerate}
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    % $\aTr{\cEvs}{\bForm}$ implies $\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm$,
    $\aTr{\cEvs}{\bForm}$ implies $(\aVal{=}\aReg\lor\RW)\limplies\bForm$, when $\aEvs\neq\emptyset$,
  \item[{\labeltext[L6]{L6)}{L6}}] 
    $\aTr{\dEvs}{\bForm}$ implies $\bForm$, when $\aEvs=\emptyset$.
  \end{enumerate}  
\end{definition}
\ref{L4} is the \emph{dependent} case,
\ref{L5} the \emph{independent} case, and
\ref{L6} the \emph{empty} case. 
 Recall the substitutions performed by $\ref{sTOP}$
(\refdef{def:pomsets-top}).
For writes, \ref{L5} is the same as \ref{L6} (since $\RW$
is $\TRUE$).  For reads, \ref{L5} is the
same as \ref{L4} (since $\RW$ is $\FALSE$).
% For reads, the dependent and independent predicate transformers are identical
% .  For writes, \ref{L5} requires
% $\aTr{\cEvs}{\bForm}$ implies $\bForm$ for the independent case.

One reading of \ref{L5} is that when satisfying a precondition $\aForm$ it is
safe to use the value of the read, as long as the action is not a write.  If
the pomset is empty, however, there is no value $\aVal$ to use.  Therefore
the best we can do is to emulate skip, as in \ref{L6}.  % In order to
% eventually arrive at a top-level pomset, this means that subsequent code must
% be independent of $\aReg$.

\begin{example}
  Revisiting \refex{ex:rrd1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \event{b}{\DR{y}{1}}{}
          \xform{x1i}{(1{=}r\lor\RW)\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{r\mid\DR{x}{0}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing sequentially:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \SEMI
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{(1{=}r\lor\RW)\limplies r\mid\DR{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  At top-level, $\ref{sTOP}$ yields:
  \begin{align*}
    \begin{gathered}[t]
      % \PR{y}{r}
      % \SEMI
      % \IF{r}\THEN s\GETS x\FI
      % \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{(1{=}r\lor\FALSE)\limplies r\mid\DR{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  The precondition of $\DRP{x}{0}$ is a tautology, as required.
\end{example}

\subsection{Full semantics for ARM}

\refdef{def:pomsets-arm} combines all of the features of
\textsection\ref{sec:q}--\ref{sec:arm}.  
% Only \ref{L5} and \ref{T3} are changed from \refdef{def:pomsets-down}.

>>>>

\section{Arm}
For simplicity, we restrict to top level parallel composition and ignore
fences\footnote{Fences are not actions in \armeight{}, which complicates the
  theorem statements.}.

\subsection{Arm executions}
\begin{definition}
  An \emph{\armeight{} execution graph}, $\aEG$, is tuple
  $(\aEvs, \labeling, {\rpoloc}, {\rlob})$ such that
  \begin{enumerate}[,label=(\textsc{a}\arabic*),ref=\textsc{a}\arabic*]
  \item $\Event\subseteq\AllEvents$ is a set of {events},
  \item $\labeling: \Event \fun \Act$ defines a {label} for each event,
  \item ${\rpoloc} : \Event\times\Event$, is a per-thread, per-location total
    order, capturing \emph{per-location program order},
  \item ${\rlob} : \Event\times\Event$, is a per-thread partial order capturing
    \emph{locally-ordered-before}, such that
    \begin{enumerate}
    \item \label{arm-lob-poloc}
      ${\rpoloc} \cup {\rlob}$ is acyclic.
    \end{enumerate}
  \end{enumerate}
\end{definition}

The definition of $\rlob$ is complex.  Comparing with our definition of
sequential composition, it is sufficient to note that $\rlob$ includes
\begin{enumerate}[label=(\textsc{l}\arabic*),ref=\textsc{l}\arabic*]
\item read-write dependencies, required by \ref{seq-kappa},
\item synchronization delay of ${\reorderra}$, required by \ref{seq-le-delays},
\item $\mSC$ access delay of ${\eqreordersc}$, required by \ref{seq-le-delays},
\item write-write and read-to-write coherence delay of ${\reorderco}$, required by \ref{seq-le-delays},
\end{enumerate}
and that $\rlob$ does \emph{not} include
\begin{enumerate}[resume,label=(\textsc{l}\arabic*),ref=\textsc{l}\arabic*]
\item \label{lob-rr} \labeltext[above]{}{lob-le}
  read-read control dependencies, required by \ref{seq-kappa},
\item \label{lob-rf}
  write-to-read order of $\rrfx$, required by \ref{seq-le-rf},
\item \label{lob-wr}
  write-to-read coherence delay of ${\reorderco}$, required by \ref{seq-le-delays}.
\end{enumerate}

\begin{definition}
  Execution $\aEG$ is
  \emph{$({\rco}, {\rrfx}, {\rgcb})$-valid}, under \emph{External Global
    Consistency} (\EGC{}) if
  \begin{enumerate}[label=(\textsc{a}\arabic*),ref=\textsc{a}\arabic*]
    \setcounter{enumi}{4}
  %\item[\eqref{arm-co}] and \eqref{arm-rf}, as for \EC,
  \item \label{arm-co}
    ${\rco} : \Event\times\Event$, is a per-location total order on
    writes, capturing \emph{coherence}, 
  \item \label{arm-rf}
    ${\rrfx} : \Event\times\Event$, is a surjective and injective
    relation on reads, capturing \emph{reads-from}, such that
    \begin{enumerate}
    \item \label{arm-match}
      if $\bEv\xrfx\aEv$ then $\labeling(\bEv) \rmatches \labeling(\aEv)$,      
    \item \label{arm-local}
      ${\rpoloc} \cup {\rco} \cup {\rrfx} \cup {\rfr}$ is acyclic,
      where $\aEv\xfr\cEv$ if %$(\exists\bEv)$
      $\aEv\xrfxinv\bEv\xco\cEv$, for some $\bEv$,
      % \stepcounter{enumi}
      % \item[] \labeltext[\textsc{a}2]{}{arm-cb}
    \end{enumerate}
    % \item ${\rpoloc} \cup {\rco} \cup {\rrfx} \cup {\rfr}$ is acyclic, where
    %   $\bEv\xfr\aEv$ if $(\exists\cEv)$ $\bEv\xrfxinv\cEv\xco\aEv$,
  \item \label{arm-gcb}
    ${\rgcb}\supseteq\PBR{{\rco}\cup{\rrfx}}$ is a linear order %, capturing \emph{globally completes before}, %$\rcb : % \Event\times\Event$
    such that 
    \begin{enumerate}%[leftmargin=0pt]
      % \item if $\bEv\xco\aEv$ then $\bEv\xgcb\aEv$,
      % \item if $\bEv\xrfx\aEv$ then $\bEv\xgcb\aEv$,
    \item \label{arm-gcb-blocks}
      if $\bEv\xrfx\aEv$ and $\labeling(\cEv) \rblocks \labeling(\aEv)$ then either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$, 
    \item \label{arm-gcb-lob}
      if $\aEv\xlob\cEv$ then either $\aEv\xgcb\cEv$ or $(\exists\bEv)$
      $\bEv\xrfx\aEv$ and $\bEv\xpoloc\aEv$ but not $\bEv\xlob\cEv$.
    \end{enumerate}
  \end{enumerate}

  Execution $\aEG$ is
  \emph{$({\rco}, {\rrfx}, {\rcb})$-valid} under \emph{External Consistency} (\EC{}) if
  \begin{enumerate}[resume,label=(\textsc{a}\arabic*),ref=\textsc{a}\arabic*]
  \item[\eqref{arm-co}] and \eqref{arm-rf}, as for \EGC,
  \item \label{arm-cb}
    ${\rcb}\supseteq\PBR{{\rco}\cup{\rlob}}$ is a linear order %, capturing \emph{completes before}, %$\rcb : % \Event\times\Event$
    such that if $\bEv\xrfx\aEv$ then either
    \begin{enumerate}%[leftmargin=0pt]
    \item \label{arm-rfe}
      % $\bEv\xcb\aEv$ and $\cEv\rblocks\aEv$ then either $\cEv\xcb\bEv$ or $\aEv\xcb\cEv$, or 
      $\bEv\xcb\aEv$ and if $\labeling(\cEv) \rblocks \labeling(\aEv)$ then either $\cEv\xcb\bEv$ or $\aEv\xcb\cEv$, or
      % $\bEv\xcb\aEv$ and $(\not\exists\cEv)$ such that $\bEv\xcb\cEv\xcb\aEv$ and $\cEv\rblocks\aEv$, or 
    \item \label{arm-rfi}
      $\bEv\xcbinv\aEv$ and $\bEv\xpoloc\aEv$ and $(\not\exists\cEv)$ $\labeling(\cEv) \rblocks \labeling(\aEv)$ and $\bEv\xpoloc\cEv\xpoloc\aEv$.
      % $\bEv\xpoloc\aEv$ and $(\not\exists\cEv)$ such that $\bEv\xpoloc\cEv\xpoloc\aEv$ and $\cEv\rblocks\aEv$.
      % \item if $\bEv\xco\aEv$ then $\bEv\xgcb\aEv$,
      % \item if $\bEv\xrfx\aEv$ then $\bEv\xgcb\aEv$,
      % \item if $\bEv\xrfx\aEv$ and $\cEv\rblocks\aEv$ then either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$,
      % \item if $\bEv\xlob\aEv$ then either $\bEv\xgcb\aEv$ or $(\exists\cEv)$
      %   $\cEv\xrfx\bEv$ and $\cEv\xpoloc\bEv$ but not $\cEv\xlob\aEv$.
    \end{enumerate}
  \end{enumerate}
\end{definition}
\cite{armed} explain \EGC{} and \EC{} using the following example.\footnote{We have
changed an address dependency in the first thread to a data dependency.}
\begin{gather*}
  \PW{x}{1}\SEMI 
  \PR{x}{r}\SEMI
  \PW{y}{r} \PAR
  \PR[\mACQ]{y}{1}\SEMI
  \PR{x}{s}
  \\
  \tag{\cmark\armeight}
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{1}}{}
      \event{b}{\DR{x}{1}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \raevent{d}{\DR[\mACQ]{y}{1}}{right=2.5em of c}
      \event{e}{\DR{x}{0}}{right=of d}
      \rfx{a}{b}
      \lob{b}{c}
      \rfx{c}{d}
      \lob{d}{e}
      \co[out=-165,in=-15]{e}[above right]{a}
    \end{tikzinline}}
\end{gather*}
\EGC{} drops $\rlob$-order in the first thread using \ref{arm-gcb-lob}, since
$\DWP{x}{1}$ is not $\rlob$-ordered before $\DWP{y}{1}$.
\begin{gather*}
  \tag{$\rgcb$}
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{1}}{}
      \event{b}{\DR{x}{1}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \raevent{d}{\DR[\mACQ]{y}{1}}{right=2.5em of c}
      \event{e}{\DR{x}{0}}{right=of d}
      \gcbz{a}{b}
      \gcbz{c}{d}
      \gcbz{d}{e}
      \gcbz[out=-165,in=-15]{e}{a}
      % \rfx{a}{b}
      % %\lob{b}{c}
      % \rfx{c}{d}
      % \lob{d}{e}
      % \co[out=-165,in=-15]{e}[above right]{a}
    \end{tikzinline}}
\end{gather*}
\EC{} drops $\rrfx$-order in the first thread using \ref{arm-rfi}.
\begin{gather*}
  \tag{$\rcb$}
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{1}}{}
      \event{b}{\DR{x}{1}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \raevent{d}{\DR[\mACQ]{y}{1}}{right=2.5em of c}
      \event{e}{\DR{x}{0}}{right=of d}
      \cbz{b}{c}
      \cbz{c}{d}
      \cbz{d}{e}
      \cbz[out=-165,in=-15]{e}{a}
      % %\rfx{a}{b}
      % \lob{b}{c}
      % \rfx{c}{d}
      % \lob{d}{e}
      % \co[out=-165,in=-15]{e}[above right]{a}
    \end{tikzinline}}
\end{gather*}

\subsection{Arm Compilation 1}

\citet{DBLP:journals/pacmpl/PodkopaevLV19} lowers to \armeight{} as follows:
Relaxed access is implemented using \LDR/\STR, non-relaxed access using
\LDAR/\STLR.  %, acquire and other fences using \DMBLD/\DMBSY.
In this section, we consider a suboptimal strategy, which lowers non-relaxed
reads to $(\DMBSY\SEMI\LDAR)$.

We do not distinguish control dependencies from address dependencies, and
therefore \ref{lob-rr} forces us to drop all dependencies between reads.  To
achieve this, we modify the definition of $\labelingForm'_2$ in
\reffig{fig:sem}.
\begin{definition}
  \label{def:semrr}
  Let $\semrr{}$ be defined as in \reffig{fig:sem}, replacing the definition
  of $\labelingForm'_2$ with:
  \begin{displaymath}
    \labelingForm'_2(\aEv)=
    \begin{cases}
      \aTr[1]{}{\labelingForm_2(\aEv)} & \text{if}\; \labeling(\aEv) \;\text{is a read}
      \\
      \aTr[1]{\Cdown{\aEv}}{\labelingForm_2(\aEv)} & \text{otherwise, where}\; \Cdown{\aEv}=\{ \cEv \mid \cEv \lt \aEv \}
    \end{cases}
    % $\Cdown{\aEv}=\{ \cEv \mid \cEv \lt \aEv \}$ if $\labeling(\aEv)$ is a write, and $\Cdown{\aEv}=\aEvs_1$, otherwise
    % \Cdown{\aEv}=
    % \begin{cases}
    %   \{ \cEv \mid \cEv \lt \aEv \} & \textif \labeling(\aEv) \;\text{is a write}
    %   \\
    %   \aEvs_1 & \textotherwise
    % \end{cases}
  \end{displaymath}
\end{definition}

\begin{theorem}
  \label{thm:ec}
  Suppose $\aEG_1$ is $({\rco_1}, {\rrfx_1}, {\rgcb_1})$-valid for $\aCmd$
  under the suboptimal lowering that maps non-relaxed reads to
  $(\DMBSY\SEMI\LDAR)$.  Then there is a top-level pomset
  $\aPS_2\in\semrr{\aCmd}$ such that $\aEvs_2=\aEvs_1$,
  $\labeling_2=\labeling_1$, ${\rrfx_2}={\rrfx_1}$, and ${\le_2}={\rgcb_1}$.

  \vspace{-.5\baselineskip}
  \begin{proof}
    First, we establish some lemmas about \armeight.

    \vspace{-.5\baselineskip}
    \begin{lemma}
      \label{lemma:fr}
      Suppose $\aEG$ is $({\rco}, {\rrfx}, {\rgcb})$-valid.  Then
      ${\rgcb}\supseteq{\rfr}$.

      \vspace{-.5\baselineskip}
      \begin{proof}
        Using the definition of ${\rfr}$ from \ref{arm-local}, we have
        $\aEv\xrfxinv\bEv\xco\cEv$, and therefore $\labeling(\cEv)$ blocks
        $\labeling(\aEv)$.    
        Applying \ref{arm-gcb-blocks}, we have that either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$.
        Since $\rgcb$ includes $\rco$, we have $\bEv\xgcb\cEv$, and therefore it
        must be that $\aEv\xgcb\cEv$.
      \end{proof}
    \end{lemma}

    \begin{lemma}
      \label{lemma:wr}
      Suppose $\aEG$ is $({\rco}, {\rrfx}, {\rgcb})$-valid and
      $\cEv\xpoloc\aEv$, where $\labeling(\cEv)\rblocks\labeling(\aEv)$.
      Then $\cEv\xgcb\aEv$.
      %$\labeling(\cEv)=\DWP{x}{}$ and $\labeling(\aEv)=\DRP{x}{}$.

      \vspace{-.5\baselineskip}
      \begin{proof}
        By way of contradiction, assume $\aEv\xgcb\cEv$.  If $\cEv\xrfx\aEv$
        then by \ref{arm-gcb} we must also have $\cEv\xgcb\aEv$,
        contradicting the assumption that $\rgcb$ is a total order.
        %
        Otherwise that there is some $\bEv\neq\cEv$ such that
        $\bEv\xrfx\aEv$, and therefore $\bEv\xgcb\aEv$.  By transitivity,
        $\bEv\xgcb\cEv$.  By the definition of $\rfr$, we have
        $\aEv\xfr\cEv$.  But this contradicts \ref{arm-local}, since
        $\cEv\xpoloc\aEv$.
        %Applying \ref{arm-gcb-blocks}, we have that either $\cEv\xgcb\bEv$ or $\aEv\xgcb\cEv$.
      \end{proof}
    \end{lemma}
    We show that all the order required in the pomset is also required by
    \armeight{}.  \ref{rf-block} holds since ${\rcb_1}$ is consistent with
    ${\rco_1}$ and ${\rfr_1}$.  As noted \ref{lob-le}, $\rlob$ includes the order
    required by \ref{seq-kappa} and \ref{seq-le-delays}.  We need only show
    that the order removed from \ref{arm-gcb-lob} can also be removed from
    the pomset.  In order for \label{arm-gcb-lob} to remove order from $\aEv$
    to $\cEv$, we must have $\bEv\xrfx\aEv$ and $\bEv\xpoloc\aEv$ but not
    $\bEv\xlob\cEv$.  Because of our suboptimal lowering, it must be that
    $\aEv$ is a relaxed read; otherwise the $\DMBSY$ would require
    $\bEv\xlob\cEv$.  Thus we know that \ref{seq-le-delays} does not require
    order from $\aEv$ to $\cEv$.  By chaining \ref{read-tau-ind} and
    \ref{write-tau}, any dependence on the read can by satisfied without
    introducing order in \ref{seq-kappa}.
  \end{proof}  
\end{theorem}

% This model compiles correctly to arm using the lowering: Relaxed access is
% implemented using \texttt{ldr}/\texttt{str}, non-relaxed reads using
% \texttt{dmb st}\SEMI\texttt{ldar}, non-relaxed writes using \texttt{stlr},
% acquire fences using \texttt{dmb}.\texttt{ld} and other fences using 
% \texttt{dmb}.\texttt{sy}.
% \begin{align*}
%   \low{\PW[\gemode\mREL]{\REF{\aReg}}{\bReg}} &= \texttt{stlr}\;\bReg,\REF{\aReg}
%   &
%   \low{\PR[\mRLX]{\REF{\aReg}}{\bReg}} &= \texttt{ldr}\;\bReg,\REF{\aReg}
%   \\
%   \low{\PW[\mRLX]{\REF{\aReg}}{\bReg}} &= \texttt{str}\;\bReg,\REF{\aReg}
%   &
%   \low{\PR[\gemode\mACQ]{\REF{\aReg}}{\bReg}} &= \texttt{dmb st}\SEMI\texttt{ldar}\;\bReg,\REF{\aReg}
% \end{align*}
